%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\pdfminorversion=4

\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{color}
\begin{document}

\title[EEG Augmentation with Diffusion Model]{Data Augmentation for EEG Motor Imagery Classification using Diffusion Model}

\author{Nutapol Soingern$^1$,Akraradet Sinsamersuk$^2$,Chaklam Silpasuwanchai$^3$}

\address{Asian Institute of Technology, School of Engineering and Technology, Data Science and Artificial Intelligence, Pathum Thani, Thailand}
\ead{nutapol1997@gmail.com, akraradets@gmail.com, chaklam@ait.asia}
\vspace{10pt}
\begin{indented}
\item[]November 2023
\end{indented}

\begin{abstract}
Motor imagery classification using electroencephalogram (EEG) signals is an important research topic that has been extensively studied in the field of brain-computer interfaces (BCIs). 
However, due to the limited amount of available data, the possibility of overfitting is a challenge, especially when using a deep-learning classifier. 
One way to address this is by performing data augmentation. 
This paper investigates the efficacy of the diffusion model as a data augmentation method for motor imagery classification. 
We evaluated the diffusion method by comparing it with commonly-used EEG data augmentation techniques namely such as Noise Addition, Fourier Transform Surrogates, Frequency Shift, and SmoothTimeMask. 
The result shows that the diffusion method outperformed other methods in terms of classification accuracy by 17.49\%. 
The Kullback-Leibler (KL) divergence is used for assessing the similarity between the training set (with and without augmentation) and validation set, thus showing the effectiveness of the diffusion approach compared to other techniques.
\end{abstract}

%
% Uncomment for keywords
\vspace{2pc}
\noindent{\it Keywords}: Motor Imagery, EEG, Brain Computer Interface, Deep Learning, Data Augmentation, Diffusion, KL divergence
%
%
% Uncomment for Submitted to journal title message
%\submitto{\JPA}
%
% Uncomment if a separate title page is required
%\maketitle
% 
% For two-column output uncomment the next line and choose [10pt] rather than [12pt] in the \documentclass declaration
%\ioptwocol
%

\input{introduction}

\input{relatedwork}

\input{methodology}

\input{result}

\input{discussion}

\section{Conclusion}
In this study, we investigated the effectiveness of data augmentation techniques in improving the accuracy of standard EEG MI classification models. 
We used five standard EEG MI classification models to classify EEG signals into left and right hand movements. 
We evaluated the performance of data augmentation techniques by comparing them with the baseline model, which was trained without any data augmentation. 
For the data augmentation techniques, we used \texttt{WaveGrad}, \texttt{Noise Addition}, \texttt{Frequency Shift}, \texttt{Fourier transform surrogates}, and \texttt{SmoothTimeMask}.

Our results showed that data augmentation techniques improved the performance of standard EEG MI classification models, with \texttt{WaveGrad} being the most effective technique. 
The accuracy of the baseline model was 42.45\%, while the accuracy of the model trained with 75\% synthetic data generated by \texttt{WaveGrad} was 51.64\%. 
The other data augmentation techniques also improved the performance of the models, with \texttt{Noise Addition} and \texttt{Frequency Shift} being the least effective techniques.

\section{Acknowledgments}

We want to thank Dr. Itthi Chatnuntawech for his valuable input to this project. 
The authors have no conflict of interests related to this publication.
And, this publication is a work under Nutapol Soingern master thesis.

\section*{References}
\bibliographystyle{iopart-num}
\bibliography{references}






\end{document}

